---
title: calculating z scores
author: ''
date: '2020-07-16'
slug: calculating-z-scores
output:
  html_document:
    keep_md: yes
---


The scale() function will center and/or scale a numeric object. Use center = TRUE to subtract the mean from the values; use scale = TRUE to divide the centered values by the standard deviation to produce z scores.

#### center variable A 
 
 > scale(A, center = TRUE, scale = FALSE)


#### generate z-scores for variable A 

> scale(A, center = TRUE, scale = TRUE)

Lets try it with EMG data

#### read in clean EMG data

Sample data from N=5 participants

```{r message=FALSE, warning=FALSE}
library(tidyverse)

emg <- read_csv("sampleEMG.csv")

glimpse(emg)
```


# Get data set up 

This dataset is long with muscle (brow,cheek) in a single variable. We want to z score each muscle separately, so easiest to make the brow and cheek data wide. 

#### Make data wide

```{r}
emg_wide <- emg %>%
  pivot_wider(names_from = "muscle", values_from = "rms")

glimpse(emg_wide)
```

Now use mutate to create new columns that contain z scores.

#### Option 1: mutate z scores manually

> mutate(z_score = (value - mean(value)) / sd(value))

```{r}
emg_z_manual <- emg_wide %>%
  mutate(Zbrow = (brow - mean(brow))/sd(brow)) %>%
  mutate(Zcheek = (cheek - mean(cheek))/sd(cheek))
```


#### Option 2: mutate z scores using scale()

```{r}
emg_z_scale <- emg_wide %>%
  mutate(Zbrow = scale(brow, center = TRUE, scale = TRUE)) %>%
  mutate(Zcheek = scale(cheek, center = TRUE, scale = TRUE))
```

#### Check

Use summary to check that the mean is now 0

```{r}
summary(emg_z_scale)
```


## EMG z scores by participant

In the context of EMG, we want to standardise scores separately for each participant. You can do that using group_by(pp_no) before the mutate. 


```{r}
emg_z_scale_bypp <- emg_wide %>%
  group_by(pp_no) %>%
  mutate(Zbrow = scale(brow, center = TRUE, scale = TRUE)) %>%
  mutate(Zcheek = scale(cheek, center = TRUE, scale = TRUE))
```


Use a filter then summary to check that it gives you the same values as before for a single participant.

```{r}
emg_z_scale_bypp %>%
  filter(pp_no == "pp1") %>%
  summary()
```


