---
title: rebecca barter purrr
author: ''
date: '2019-08-20'
slug: rebecca-barter-purrr
categories: []
draft: TRUE
tags: []
---

Rebecca Barter spoke to RLadies Sydney this week about `purrr` and I think I've finally got it!

I even kinda understand what a list is now. Lists are objects that contain more than 1 type of object (vectors, numbers, dataframes) within them. 

```{r include=FALSE}
library(tidyverse)
```

```{r}
my_first_list <- list(my_number = 5,
                      my_vector = c("a", "b", "c"),
                      my_dataframe = data.frame(a = 1:3, b = c("q", "b", "z"), c = c("bananas", "are", "so very great")))
my_first_list
```

Dataframes are special kinds of lists in each object is the same length. 

#### Make a function that adds 10

```{r}
addTen <- function(.x) {
  return(.x + 10)
}
```

Test the function on a single item
```{r}
addTen(47)
```

Iterate across a vector

```{r}
map(.x = c(1, 4, 7), 
    .f = addTen)
```

You don't need to specify arguments. Just remember that map expectss the first argument to be the data you want to iterate across and the second argument is the function

```{r}
map(c(1, 4, 7), addTen)
```

The map function will run on a list or dataframe. You can use the map variants (like map_dbl) to specify what the product should me. 
This one returns a double
```{r}
map_dbl(c(1, 4, 7), addTen)
```
This one returns a character
```{r}
map_chr(c(1, 4, 7), addTen)
```

To return a df, return both the old and new. 
```{r}
map_df(c(1, 4, 7), function(.x) {
  return(data.frame(old_number = .x, 
                    new_number = addTen(.x)))
})
```

The modify function is similar to map but returns the same type as the input.  

Vector
```{r}
modify(c(1, 4, 7), addTen)
```

List
```{r}
modify(list(1, 4, 7), addTen)
```
Dataframe
```{r}
modify(data.frame(1, 4, 7), addTen)
```

Modify if applies the function to items that meet the criteria. This will apply the add ten function to items greater than 5. Here .x i the data, .p is the predicate function, and .f is the function. 
```{r}
modify_if(.x = list(1, 4, 7), 
          .p = function(x) x > 5,
          .f = addTen)
```

The tilde-dot short hand ~.

This is the same as ....
```{r}
function(x) {
  x + 10
}
```

this...
```{r}
~{.x + 10}
```

you can define the function within the map, which makes everything shorter

```{r}
map_dbl(c(1, 4, 7), ~{.x + 10})
```

EXAMPLES

```{r}
# to download the data directly:
gapminder_orig <- read.csv("https://raw.githubusercontent.com/swcarpentry/r-novice-gapminder/gh-pages/_episodes_rmd/data/gapminder-FiveYearData.csv")
# define a copy of the original dataset that we will clean and play with 
gapminder <- gapminder_orig
```

Get the class for each column
```{r}
gapminder %>% map_chr(class)
```
Get the number of distinct values for each column
```{r}
gapminder %>% map_dbl(n_distinct)
```
Make a new data frame of the number of distinct values and the class of each column
```{r}
gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))))
```

Oops lost the variable names, need to include id

```{r}
gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "variable")
```

### thinking through map

Make it work on 1 instance, then insert it into a map

For example, define .x as the first element of gapminder
```{r}
.x <- gapminder %>% pluck(1)
# look at the first 6 rows
head(.x)
```

create a dataframe containing the distinct values and class for this instance

```{r}
data.frame(n_distinct = n_distinct(.x),
           class = class(.x))
```

THen pop that into a map 

```{r}
gapminder %>% map_df(~(data.frame(n_distinct = n_distinct(.x),
                                  class = class(.x))),
                     .id = "variable")
```


Map across several objects with map2

map2(.x = object1, # the first object to iterate over
     .y = object2, # the second object to iterate over
     .f = plotFunction(.x, .y))

Get vector of continents and years

```{r}
continent_year <- gapminder %>% distinct(continent, year)
continent_year
```

```{r}
# extract the continent and year pairs as separate vectors
continents <- continent_year %>% pull(continent) %>% as.character
years <- continent_year %>% pull(year)
```

```{r}
# try to figure out the code for the first example
.x <- continents[1]
.y <- years[1]
# make a scatterplot of GDP vs life expectancy in all Asian countries for 1952
gapminder %>% 
  filter(continent == .x,
         year == .y) %>%
  ggplot() +
  geom_point(aes(x = gdpPercap, y = lifeExp)) +
  ggtitle(glue::glue(.x, " ", .y))
```

Then put the code into map, making a plot list

```{r}
plot_list <- map2(.x = continents, 
                  .y = years, 
                  .f = ~{
                    gapminder %>% 
                      filter(continent == .x,
                             year == .y) %>%
                      ggplot() +
                      geom_point(aes(x = gdpPercap, y = lifeExp)) +
                      ggtitle(glue::glue(.x, " ", .y))
                  })
```

Pick a few entries to make sure they make sense

```{r}
plot_list[[1]]

plot_list[[22]]
```

Tibbles = tidyverse data frames

AND columns can be lists rather than just vectors. You can have a nested tibble
```{r}
gapminder_nested <- gapminder %>% 
  group_by(continent) %>% 
  nest()
gapminder_nested
```

```{r}
gapminder_nested$data[[1]]
```
dplyr pluck will pull the first entry, which is asia
```{r}
gapminder_nested %>% 
  # extract the first entry from the data column
  pluck("data", 1)
```

pluck 5 is oceania
```{r}
gapminder_nested %>% pluck("data", 5)
```

why would you do this? you can use dplyr on super complex structures

```{r}
tibble(vec_col = 1:10) %>%
  mutate(vec_sum = sum(vec_col))
```

