---
title: lubridate math with dates
author: ''
date: '2020-10-19'
slug: lubridate-math-with-dates
categories: []
draft: TRUE
tags: []
---

So far in this `lubridate` journey I have learned ...

1. how to pull date components (day, month, year) from date data in R and use them within dplyr and ggplot functions
2. how to use `dmy()` and friends to convert character formats into date formats
3. that whether or not R recognises your dates as dates is a bit of a crap shoot
4. that it is best not to give your participants free reign with date entry, but if you do that `parse_date_time()` does a pretty good job 

The one last thing that `lubridate` is useful for is doing math with dates. 

# Time spans in R

For this one I looked up the [R4DS book](https://r4ds.had.co.nz/dates-and-times.html) as a reference guide. R thinks about 3 different types of time span. 

1. durations - these are a length of time in seconds
2. period - these are a length of time in useful to think about units (i.e. weeks or months)
3. intervals - these are a length of time between a start and end point

I'm going to be honest, duration I get, but the distinction between period and interval was initially lost on me. 

Lets work through the exercises. 

```{r}
library(lubridate)
```


# Durations

When you subtract two dates in R you get a difftime object, which can be expressed as seconds, minutes, hours, days or weeks. 

```{r}
j_age <- today() - dmy(30061978)

j_age
```

To keep things consistent, lubridate specifies durations, which are always expressed as seconds. 

```{r}
as.duration(j_age)
```

Use dseconds(), dminutes(), dhours(), ddays(), dweeks(), dyears() to get how long a duration is, that is how many seconds are in X number of seconds, minutes, hours, days, weeks or years. 

```{r}

# d-functions create durations in seconds
dseconds(30)

dminutes(30)

dhours(30)

ddays(30)

dweeks(30)

dyears(30)
```

By reducing everything to the same units, you can more easily do math with durations.

```{r}
2 * dyears(1)

dyears(1) + dweeks(12) + dhours(15)
```

You can add and subtract durations. 

```{r}
today() + ddays(1)

today() - dyears(1)
```

But sometimes timezones and daylight savings make for complications when you add days as durations. Solution = periods...

# Periods

Periods are time spans that have more usable units, that is year, month, day, hour, min, second.  (y,m,d,H,M,S)


```{r}
seconds(30)

minutes(30)

hours(30)

days(30)

weeks(30)

years(30)
```

